---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Here is the code for loading data
```{r loaddata, echo=TRUE}
Mydata <- read.csv(file = "C:/Users/yzhang/Documents/GitHub/RepData_PeerAssessment1/activity/activity.csv", header = TRUE, sep = ",")
```
## What is mean total number of steps taken per day?
Here is the code to calculate the total steps
```{r CalculateStepsperDay, echo=TRUE}
totalStepbyDay <- tapply(Mydata$steps, Mydata$date, sum, na.rm = TRUE)
```
Below is the code to generate the histogram
```{r Createhistogram, echo=TRUE}
hist(totalStepbyDay, breaks = 10)
```
Below is the code for calculating the mean and its output
```{r CalculateMean, echo=TRUE}
Mymean <- mean(totalStepbyDay)
print(Mymean, type = "html")
```
Below is the code for calculating the median and its output
```{r CalculateMedian, echo=TRUE}
Mymedian <- median(totalStepbyDay)
print(Mymedian, type = "html")
```

## What is the average daily activity pattern?
Below is the code for calculate the average steps grouping by intervals.
```{r CalculateAverageSteps, echo=TRUE}
library(dplyr)
StepsbyInterval <- Mydata %>%
  group_by(interval) %>%
  summarise(average_steps = mean(steps, na.rm = TRUE))
```
Plot the data
```{r PlotLineChart, echo = TRUE}
with(StepsbyInterval, plot(interval, average_steps, type = "l", xlab = "intervals", ylab = "average steps", main = "Average Daily Activity Pattern"))
```
Report which 5-minute interval contains the maximum number of steps.
```{r ReportMaxiumStepsbyInterval, echo=TRUE}
MaxiumStepsbyInterval <- StepsbyInterval[StepsbyInterval$average_steps == max(StepsbyInterval$average_steps), ]
print(MaxiumStepsbyInterval, type = "html")
```

## Imputing missing values




## Are there differences in activity patterns between weekdays and weekends?
